---
title: "Demographics and Baseline Characteristics (DMT01)"
editor: source
format: 
  html:
    page-layout: full
    code-fold: true
    code-summary: "Show the code"
---

## Building a DMT01 table using the {chevron} package


```{r}
#| message: false
#| code-summary: "Show the code"
library(chevron)

# Create a table using the chevron package 
dmt01 <- chevron::run(dmt01, syn_data)
dmt01

```

We can use {rtables} functionality to flatten this table into a data frame.
```{r}
results <- rtables::as_result_df(dmt01)
results [, -c(1:4,6:8)]
```


## Building an DMT01 table using the {cards} package

In this example below we create the same demographics table using the {gtsummary} package, with the same data set provided by the {chevron} package.

```{r}
#| code-summary: "Show the code"
library(gtsummary)

tbl <-
  syn_data$adsl|>
  gtsummary::tbl_summary(
  by = ARM,
  statistic = list(
    all_continuous() ~ "{mean} ({sd})",
    all_categorical() ~ "{n} / {N} ({p}%)"
  ),
  missing = "no",
  include = c("AGE", "AGEGR1", "ARM", "SEX", "ETHNIC", "RACE")
)

ard <- tbl |> gather_ard()

ard$tbl_summary
```

## Comparing Outputs
Upon comparing the two outputs, it's evident that both methods yield identical values, enabling their use for Quality Control (QC) purposes.
