---
title: "Demographics and Baseline Characteristics (DMT01)"
editor: source
format: 
  html:
    page-layout: full
    code-fold: true
    code-summary: "Show the code"
---

## Building a DMT01 table using the {chevron} package


```{r}
#| message: false
#| code-summary: "Show the code"
library(chevron)

# Create a table using the chevron package 
dmt01 <- chevron::run(dmt01, syn_data)
dmt01

```

We can use {rtables} functionality to flatten this table into a data frame.
```{r}
results <- rtables::as_result_df(dmt01)
results [, -c(1:4,6:8)]
```


## Building an DMT01 table using the {cards} package

In this example below we create the same demographics table using the {gtsummary} package, with the same data set provided by the {chevron} package.

```{r}
#| code-summary: "Show the code"
library(cards)

# build ARDs that calculate relevant statistics for continuous and categorical variables.

ards <- bind_ard(ard_continuous(syn_data$adsl, variables = c(AGE), by = ARM),
                 ard_categorical(syn_data$adsl, variables = c(AGEGR1, SEX, ETHNIC, RACE), by = ARM))
```

## Comparing Outputs

```{r}
# Equality in means and sd

mean_tbl <- results |>
  dplyr::filter(label_name %in% "Mean (SD)") |>
  dplyr::select(9:11) |>
  dplyr::mutate(across(everything(), ~purrr::map_chr(.x, ~paste(.x, collapse = ", "))))

mean_ard <- ards %>%
  dplyr::filter(stat_name %in% c("mean", "sd")) %>%
  dplyr::summarise(
    "Mean(SD)" = paste(stat[stat_name == "mean"], "(", stat[stat_name == "sd"], ")", sep = ""))

mean_tbl
```
```{r}
mean_ard
```

```{r}
# Equality in n, N and percent

cat_tbl <- results |>
  dplyr::filter(avar_name %in% "RACE") |>
  dplyr::select(c(3,9:11)) |>
  dplyr::mutate(across(everything(), ~purrr::map_chr(.x, ~paste(.x, collapse = ", "))))

cat_tbl
```

```{r}
cat_ard <- ards %>%
  dplyr::filter(variable %in% c("RACE")) |>
  dplyr::filter(stat_name %in% c("n", "N", "p")) 

cat_ard
```

Upon comparing the two outputs, it's evident that both methods yield identical values, enabling their use for Quality Control (QC) purposes.
