---
title: "QCing DMT01"
editor: source
format: 
  html:
    page-layout: full
    code-fold: true
    code-summary: "Show the code"
    code-overflow: scroll 
    df-print: kable
---

#### Generate a table using {chevron}

```{r}
#| message: false
#| code-summary: "Show the code"
library(chevron)

# Create a table using the chevron package 
dmt01 <- chevron::run(dmt01, syn_data)
dmt01

```

#### Flatten the table into a data.frame

A rtables based output can be flattened into a data frame using the `as_results_df()` function from the rtables package. The `make_ard` argument set to `TRUE`, will format the data similar to the output generated by the {cards} package.

```{r}
results <- rtables::as_result_df(dmt01, make_ard = TRUE, expand_colnames = TRUE)
results[1:6, -c(1:3)]
```

#### Create a comparable ARD

Using the {cards} package, we stack the functions `ard_continuous()` for the continuous variables and `ard_categorical()` for categorical variables. The default statistics calculated for each of these data types are included - these can be adapted for bespoke analyses.

If any variable manipulation was done to the data prior to running the citril/chevron/tern/rtables commands, we suggest supplying the same data to these ARD functions, or running the same pre-processing steps to your data prior to creating ARDs to ensure variable names/levels match.

```{r}
#| code-fold: false
#| code-summary: "Show the code"
library(cards)

# build ARDs that calculate relevant statistics for continuous and categorical variables.

ards <- ard_stack(syn_data$adsl, ard_continuous(variables = c(AGE), statistic = ~ continuous_summary_fns(c("N", "mean", "sd", "median", "min", "max"))
  ),
                 ard_categorical(variables = c(AGEGR1, SEX, ETHNIC, RACE)),
                 .by = "ARM",
                 .overall = TRUE)

ards [1:6, -c(1,9:11)]
```

#### Statistics comparison

##### rtables reformat

In order to compare the two data.frames programatically, some identifying variables must align to be used as "key columns". Below are some data wrangling steps used to match the statistics for comparison. Note the rtables output:

```{r}
tail(results[, -c(1:3)])
```

The `variable_level` leads with the statistic name, followed by "." and then the actual level that matches the `variable_level` in the ARD object. We will mutate the level to match the level values in the ard object. Similarly, the total number of observations for a group is labelled "N" in the ARD object, while it is named "n" in the rtables object. The following manipulations are completed below:

1.  Set the `variable_level` to NA in the rtables result for a continuous data summary (as variable levels don't apply and is NULL in the ARD object).
2.  Remove the stat_name (ie. "count") before the variable level.
3.  Convert the "n":"N" and "count":"n".
4.  Recode the variable AAGE to AGE.
5.  Recode the `SEX` variable Male:M, Female:F. Note: if you are mapping variables in {citril}, these changes would already be captured.

```{r}
# rework the rtables output to match ----

reformat <- results |>
  dplyr::mutate(
    variable_level = dplyr::case_when(
      variable_level %in% c("mean_sd", "median", "range", "n") ~ NA_character_,
      TRUE ~ variable_level
    ),
    variable_level = sub("^[^.]*\\.", "", variable_level),  # use variable_label
    stat_name = dplyr::recode(stat_name, "n" = "N", "count" = "n"),
    variable = dplyr::recode(variable, "AAGE" = "AGE"),
    variable_level = dplyr::recode(variable_level, "Male" = "M", "Female" = "F")
  )

```

Notice, in the rtables output, any missing or o observations are pruned from the resulting table. ARDs capture statistics, even for 0 observations (and therefore has more rows than the rtables result). We can remove those rows before comparison to isolate for rows that are different between the two results *and* contain comparable stats.
```{r}

reformat <- reformat |>
  dplyr::filter(stat != 0,
               !(variable_label == "n" & variable_level == "n"))

```


##### ARD reformat

A reformatting step is necessary for the ARD output to complete the comparison. Note that most of the columns in the {cards} objects are a list (for example, the level of the variable followed by the levels available for the factor).

```{r}
ards$group1_level[1]
```

We will pull the first string from the list for the dataframe (group1_level, variable_level). We'll also add the string "ARM" to any NULL observations in the `group1` column.

```{r}
ards2 <- ards |>
  dplyr::mutate(
    group1_level = purrr::map_chr(group1_level, ~ ifelse(length(.x) > 0, as.character(.x[[1]]), NA_character_)),
    variable_level = purrr::map_chr(variable_level, ~ ifelse(length(.x) > 0, as.character(.x[[1]]), NA_character_)),
    stat = purrr::map_dbl(stat, ~ ifelse(length(.x) > 0, as.numeric(.x[[1]]), NA_character_)),
    group1 = dplyr::coalesce(group1, "ARM"),
    group1_level = dplyr::coalesce(group1_level, "All Patients")
  ) |>
  dplyr::select(-c(9:11)) |>
  as.data.frame()
```

#### Compare programmatically

Here we propose using the {diffdf} package to compare the statistics produced by the two table engines. {diffdf} is designed to compare two data frames and report any differences/inconsistencies to the user.

Note: We'll remove the columns that are not shared between the two results to avoid messaging for known differences (ie. fmt_fn, stat_string)

```{r}
library(diffdf)

diffdf(reformat[,!(names(reformat) %in% c("variable_label", "stat_string"))], ards2[,!(names(ards2) %in% c("fmt_fn", "context", "stat_label"))], keys = c("group1_level","group1", "variable", "variable_level", "stat_name"), suppress_warnings = TRUE)

```

If dplyr is readily available, the `anti_join()` function can be used to isolate the differences between the two dataframes for closer analysis.

```{r}
library(dplyr)

differences <- ards2 %>% #start with the data.frames which contain greater number of rows.
  dplyr::anti_join(reformat, by = c("group1_level","group1", "variable", "variable_level", "stat_name"))  # Compare based on key columns

differences

```
