---
title: "QCing Demographics and Baseline Characteristics (DMT01)"
editor: source
format: 
  html:
    page-layout: full
    code-fold: true
    code-summary: "Show the code"
    code-overflow: scroll 
    df-print: kable
---

## Demographics and Baseline Characteristics (DMT01) QC Workflow

### 1. Generate a table using {chevron}

```{r}
#| message: false
#| code-summary: "Show the code"
library(chevron)

# Create a table using the chevron package
tlg_dmt01 <- chevron::dmt01_main(syn_data, summaryvars = c("AGE", "AGEGR1", "SEX", "ETHNIC", "RACE"))
head(tlg_dmt01, n = 15)
```

### 2. Flatten the table into a data.frame

A {rtables} based output can be flattened into a data.frame using the `as_results_df()` function from the {rtables} package. 
The `make_ard` argument set to `TRUE`, will format the data similar to the output generated by the {cards} package.

```{r, warning=FALSE}
#| code-fold: false
#| code-summary: "Show the code"
rtables_result <- rtables::as_result_df(tlg_dmt01, make_ard = TRUE, add_tbl_str_decimals = FALSE)
rtables_result[1:6, -c(2:4)]
```

### 3. Create a comparable ARD

Using the {cards} package, we stack the functions `ard_continuous()` for the continuous variables and `ard_categorical()` for categorical variables. The default statistics calculated for each of these data types are included - these can be adapted for bespoke analyses.

If any variable manipulation was done to the data prior to running the {citril}/{chevron}/{tern}/{rtables} commands, we suggest supplying the same data to these ARD functions, or running the same pre-processing steps to your data prior to creating ARDs to ensure variable names/levels match.

Note: when a `by` variable is specified in the  `ard_stack()` call, a univariate tabluation is returned. We will have to remove those observations later in the workflow to create a 1:1 match. 

```{r}
#| code-fold: false
#| code-summary: "Show the code"
library(cards)

# build ARDs that calculate relevant statistics for continuous and categorical variables.
ard_result <-
  ard_stack(
    syn_data$adsl,
    ard_continuous(
      variables = c(AGE),
      statistic = ~ continuous_summary_fns(c("N", "mean", "sd", "median", "min", "max"))
    ),
    ard_categorical(variables = c(AGEGR1, SEX, ETHNIC, RACE), statistic = everything() ~ c("n", "p")),
    ard_missing(variables = c(AGEGR1, SEX, ETHNIC, RACE), statistic = everything() ~ c("N_obs")),
    .by = "ARM",
    .overall = TRUE
  )

ard_result[1:6, -c(1, 9:11)]
```

### 4.Statistics comparison

#### A. {rtables} reformat
In order to compare the two data.frames programatically, some identifying variables must align to be used as "key columns". Below are some data wrangling steps used to match the statistics for comparison. Note the {rtables} output:

```{r}
tail(rtables_result[, -c(1:2)])
```

The `variable_level` leads with the statistic name, followed by `"."` and then the actual level that matches the `variable_level` in the ARD object. We will mutate the level to match the level values in the ARD object. Similarly, the total number of observations for a group is labelled `"N"` in the ARD object, while it is named `"n"` in the {rtables} object. The following manipulations are completed below:

1.  Set the `variable_level` to `NA` in the {rtables} result for a continuous data summary (as variable levels don't apply and is `NULL` in the ARD object).
2.  Remove the stat_name (ie. "count") before the variable level.
3.  Convert the "n":"N" and "count":"n".
4.  Remove columns we know won't be in the ARD data.frame for simplicity (ie.variable label)

```{r}
rtables_result <- rtables_result |>
  dplyr::mutate(
    variable_level = dplyr::case_when(
      variable == "AGE" & variable_level %in% c("mean_sd", "median", "range", "n") ~ NA_character_,
      TRUE ~ variable_level
    ),
    variable_level = sub("^[^.]*\\.", "", variable_level), # use variable_label
    stat_name = dplyr::recode(stat_name, "n" = "N", "count" = "n")
  ) |>
  dplyr::select(-c("variable_label"))

head(rtables_result, n = 5)
```

#### B. ARD reformat
A reformatting step is necessary for the ARD output to complete the comparison. Note that many of the columns in the {cards} objects are a list (for example, the level of the variable followed by the levels available for the factor).

```{r}
ard_result$group1_level[1]
```

We will pull the first string from the lis (`group1_level`, `variable_level`). We'll also add the string "ARM" to any `NULL` observations in the `group1` column to match the {rtables} result.

```{r}
ard_result <- ard_result |>
  dplyr::mutate(
    group1_level = purrr::map_chr(group1_level, ~ ifelse(length(.x) > 0, as.character(.x[[1]]), NA_real_)),
    variable_level = purrr::map_chr(variable_level, ~ ifelse(length(.x) > 0, as.character(.x[[1]]), NA_real_)),
    stat = purrr::map_dbl(stat, ~ ifelse(length(.x) > 0, as.numeric(.x[[1]]), NA_real_)),
    group1 = dplyr::coalesce(group1, "ARM"),
    group1_level = dplyr::coalesce(group1_level, "All Patients"),
    stat_name = dplyr::recode(stat_name, "N_obs" = "N")
  ) |>
  dplyr::select(-c("context", "fmt_fn", "stat_label", "warning", "error"))

head(ard_result, n = 5)
print("ARD dimensions")
dim(ard_result)
print("rtables Dimensions")
dim(rtables_result)
```

Note that the ARD result is larger than the {rtables} result. 
When using `ard_stack` with a listed `by` variable, a univariate analysis is run for that variable. 
Here, it is `"ARM"`.  We can remove those statistics as they are not in the `rtables_result`.

```{r}
#| code-fold: false
#| code-summary: "Show the code"
ard_result <- ard_result |>
  dplyr::filter(
    !((variable == "ARM")) | is.na(variable_level)
  ) |>
  dplyr::mutate(variable_level = dplyr::if_else(stat_name == "N" & is.na(variable_level) & variable != "AGE", "n", variable_level))
```

#### C. Compare programmatically
Here we propose using the {diffdf} package to compare the statistics produced by the two table engines. {diffdf} is designed to compare two data.frames and report any differences/inconsistencies to the user.

```{r}
#| code-fold: false
#| code-summary: "Show the code"
diffdf::diffdf(rtables_result,
  ard_result,
  keys = c("group1_level", "group1", "variable", "variable_level", "stat_name"),
  suppress_warnings = TRUE
)
```

The `anti_join()` function from the {dplyr} package can be used to isolate the differences between the two data.frames for closer analysis.

```{r}
#| code-fold: false
#| code-summary: "Show the code"
#| message: false
differences <- ard_result |> # start with the data.frames which contain greater number of rows.
  dplyr::anti_join(rtables_result,
    # Compare based on key columns
    by = c("group1_level", "group1", "variable", "variable_level", "stat_name")
  )
```
