---
title: "TTET01 - Time-to-Event"
editor: source
format: 
  html:
    page-layout: full
    code-fold: false
    code-summary: "Show the code"
    code-overflow: scroll 
    df-print: kable
---

# Time-to-event Summary (TTET01) QC Workflow

```{r}
#| message: false
# load libraries
library(cards)
library(cardx)
library(dplyr)
library(survival)
library(broom.helpers)
```

## 1. Generate a table using {chevron}

```{r}
#| message: false
#| code-fold: true
#| code-summary: "Show the code"
# Create a table using the chevron package
proc_data <- dunlin::log_filter(chevron::syn_data, PARAMCD == "PFS", "adtte")
tlg_ttet01 <- chevron::run(chevron::ttet01, proc_data)
head(tlg_ttet01, n = 15)
```

## 2. Flatten the table into a data.frame

A rtables based output can be flattened into a data frame using the `as_results_df()` function from the rtables package. The `make_ard` argument set to `TRUE`, will format the data similar to the output generated by the {cards} package.

```{r, warning=FALSE}
rtables_results <- rtables::as_result_df(tlg_ttet01, make_ard = TRUE, add_tbl_str_decimals = FALSE)
head(rtables_results, n = 5)
```

## 3. Create comparable ARDs

Using the {cards} package, we stack the functions `ard_continuous()` for the continuous variables and `ard_categorical()` for categorical variables. The default statistics calculated for each of these data types are included - these can be adapted for bespoke analyses.

If any variable manipulation was done to the data prior to running the citril/chevron/tern/rtables commands, we suggest supplying the same data to these ARD functions, or running the same pre-processing steps to your data prior to creating ARDs to ensure variable names/levels match.

Note: when a 'by' variable is specified in the `ard_stack()` call, a univariate tabulation is returned. We will have to remove those observations later in the workflow to create a 1:1 match.

```{r}
#| message: false
# Data Pre-processing
adtte <- proc_data$adtte |>
  dplyr::mutate(
    AVAL = (AVAL / 30.4375), # convert days to months, this is the same calculation chevron uses.
    is_event = CNSR == 0, # chevron coding of events
    is_not_event = CNSR == 1
  )

# ----- ARDS -----
# Calculate Event Rates
ard_event_rates <- cards::bind_ard(
  # Patient Event Rates
  ard_categorical(adtte, strata = c(ARM), variable = c(is_event), statistic = everything() ~ c("n", "p")),
  # Earliest Contributing Event
  ard_categorical(adtte |> dplyr::filter(is_event == TRUE), strata = c(ARM), variable = c(EVNTDESC), statistic = everything() ~ c("n"))
) |>
  apply_fmt_fn() |>
  unlist_ard_columns()

# Calculate Time to Event
ard_tte <- cards::bind_ard(
  # calculate median
  survfit(Surv(AVAL, is_event) ~ ARM, data = adtte, conf.type = "plain") |>
    ard_survival_survfit(probs = 0.5),
  # calculate quantiles
  survfit(Surv(AVAL, is_event) ~ ARM, data = adtte, conf.type = "plain") |>
    ard_survival_survfit(probs = c(0.25, 0.75)) |>
    filter(stat_name == "estimate"),
  # calculate range
  ard_continuous(adtte, by = c(ARM), variable = c(AVAL), statistic = ~ continuous_summary_fns(c("min", "max")))
) |>
  apply_fmt_fn() |>
  unlist_ard_columns(fct_as_chr = TRUE)

# Calculate Unstratified coxph
# set the reference arm
ref_arm <- "A: Drug X"

ard_unstrat_coxph <- ard_pairwise(
  adtte,
  variable = ARM,
  .f = \(df) {
    coxph(Surv(AVAL, is_event) ~ ARM, data = df, ties = "efron") |>
      ard_regression(exponentiate = TRUE) |>
      dplyr::filter(stat_name %in% c("estimate", "conf.low", "conf.high") & variable_level != ref_arm)
  },
  include = ref_arm
) |>
  bind_ard() |>
  apply_fmt_fn() |>
  unlist_ard_columns()

# calculate the pvalue
ard_pval <- ard_pairwise(
  adtte,
  variable = ARM,
  .f = \(df) {
    ard_survival_survdiff(survival::Surv(AVAL, is_event) ~ ARM, data = df) |>
      dplyr::filter(stat_name %in% "p.value")
  },
  include = ref_arm
) # this produces a list of data frames

ard_pval <- lapply(ard_pval, unlist_ard_columns)
ard_pval <- do.call(rbind, ard_pval)

# 6 Months Risk Analysis
ard_surv_time <- survfit(Surv(AVAL, is_event) ~ ARM, adtte, conf.int = 0.95, conf.type = "plain") |>
  ard_survival_survfit(time = 6) |>
  dplyr::filter(stat_name != "std.error") |>
  apply_fmt_fn() |>
  unlist_ard_columns()
```

### Note about survival surf_diff tables

Survival Time Difference estimates are calculated usually calculated by the treatment - the control group. It's been noted that {tern}'s `surv_timepoint` function and {cardx}'s `ard_survival_surfit_diff` function calculate this in different orders - resulting in statistics where the absolute value is identical but not the direction. If this is found in your data - you'll need to relevel the factor for the `by` variable, such that the treatment group is ordered *before* the control group. Below, we reorder the level so "B: Placebo" comes before the control group, "A: Drug X". This is repeated so "C: Combination" comes before "A: Drug X" before computing the ARD.

```{r}
# difference in event free rate
# relevel the ARM variable to match the calculation done in tern
adtte$ARM <- forcats::fct_relevel(adtte$ARM, "B: Placebo", after = 0)
ard_surv_time_diff_1 <- survfit(Surv(AVAL, is_event) ~ ARM, data = adtte, conf.int = 0.95, conf.type = "plain") |>
  ard_survival_survfit_diff(times = 6) |>
  filter(stat_name %in% c("estimate", "conf.low", "conf.high", "p.value")) |>
  apply_fmt_fn() |>
  unlist_ard_columns() |>
  filter(group1_level == "A: Drug X") |>
  mutate(group1_level = "B: Placebo")


adtte$ARM <- forcats::fct_relevel(adtte$ARM, "B: Placebo", after = 3)
ard_surv_time_diff_2 <- survfit(Surv(AVAL, is_event) ~ ARM, data = adtte, conf.int = 0.95, conf.type = "plain") |>
  ard_survival_survfit_diff(times = 6) |>
  filter(stat_name %in% c("estimate", "conf.low", "conf.high", "p.value")) |>
  apply_fmt_fn() |>
  unlist_ard_columns() |>
  filter(group1_level == "C: Combination")

ard_surv_time_diff <- bind_rows(ard_surv_time_diff_1, ard_surv_time_diff_2)

head(ard_surv_time_diff, n = 5)
```

## 4.Statistics comparison

To ensure the {rtables} and ARD results and compatiable for comparison, some reformatting will need to be applied and may be different for each sub-table. Below is a helper function created to apply a row transformation for duplicate `stat_name`in the {rtables} result. For example, two rows may be called "rate_ci" where one row is the lower bound of the confidence interval and the other is the upper bound. This function transforms these rows using the new `stat_name` values provided in the `values` parameter.

```{r}
# function to apply row-based transformation
apply_row_transformation <- function(data, condition, values) {
  data |>
    group_by(group = (variable_level == condition)) |>
    mutate(
      row_num = row_number(),
      variable_level = case_when(
        group & row_number() %% 2 == 1 ~ values[1], # Odd row in condition group
        group & row_number() %% 2 == 0 ~ values[2], # Even row in condition group
        TRUE ~ variable_level # Keep original value
      )
    ) |>
    ungroup() |>
    select(-group, -row_num) # Remove temporary columns
}
```

### Event Rate Table

For {rtables}: 1. Filter for event rate statistics. 2. Rename the grouping variables to match ARD 3. Recode "count" to "n" and "fraction" to "p" to match ARD

For ARD: 1. recode variable to match {rtables}

```{r}
# {rtables}
rtables_event_rate <- rtables_results |>
  filter(variable_label == "Patients with event (%)" | variable_label == "Patients without event (%)" | group1_level == "Patients with event (%)") |>
  dplyr::rename(
    group1 = group2,
    group1_level = group2_level,
    # rename other vars to avoid error
    group2 = group1,
    group2_level = group1_level
  ) |>
  dplyr::select(c(group1, group1_level, variable, stat_name, stat)) |>
  dplyr::mutate(
    stat_name = dplyr::recode(stat_name, "count" = "n"),
    stat_name = dplyr::recode(stat_name, "fraction" = "p")
  )

# {cards}
ard_event_rates <- ard_event_rates |>
  mutate(
    variable = recode(variable, "is_event" = "IS_EVENT"),
    variable = ifelse(variable_level == FALSE & variable != "EVNTDESC", "IS_NOT_EVENT", variable)
  ) |>
  dplyr::select(c("group1", "group1_level", "variable", "stat_name", "stat"))

# compare stats
diffdf::diffdf(
  rtables_event_rate,
  ard_event_rates,
  keys = c("group1", "group1_level", "variable", "stat_name", "stat"),
  suppress_warnings = TRUE
)
```

### Time to Event Table

For {rtables}: 1. Filter for event rate statistics. 2. Rename the grouping variables to match ARD 3. Apply row transformations to differentiate bounds of stats 4. Recode "median" to "estimate" to match ARD

```{r}
# {rtables}
rtables_tte <- rtables_results |>
  filter(variable == "time_to_event") |>
  rename(
    group1 = group2,
    group1_level = group2_level,
    group2 = group1,
    group2_level = group1_level
  ) |>
  apply_row_transformation("95% CI", c("conf.low", "conf.high")) |>
  apply_row_transformation("Range", c("min", "max")) |>
  apply_row_transformation("25% and 75%-ile", c("0.25", "0.75")) |>
  mutate(variable_level = recode(variable_level, "Median" = "estimate")) |>
  dplyr::select(c(group1, group1_level, variable_level, stat)) |>
  rename(stat_name = variable_level)

# {cards}
ard_tte <- ard_tte |>
  mutate(
    stat_name = if_else(
      variable_level %in% c("0.25", "0.75") & stat_name == "estimate",
      as.character(variable_level), # Convert to character if true
      stat_name # Keep original stat_name if false
    )
  ) |>
  dplyr::select(c("group1", "group1_level", "stat_name", "stat"))

# compare stats
diffdf::diffdf(
  rtables_tte,
  ard_tte,
  keys = c("group1", "group1_level", "stat_name", "stat"),
  suppress_warnings = TRUE
)
```

### Unstratified Pairwise Coxph

For {rtables}: 1. Filter for event rate statistics. 2. Rename the grouping variables to match ARD 3. Apply row transformations to differentiate bounds of stats 4. Recode "hr" to "estimate" to match ARD

For ARD: 1. Recode variable to match {rtables} 2. Set attributes to `NULL`

```{r}
# {rtables}
rtables_unstrat_coxph <- rtables_results |>
  filter(variable == "coxph_unstratified" & !is.na(stat) & variable_level != "pvalue") |>
  select("group2", "group2_level", "variable_level", "stat") |>
  apply_row_transformation("hr_ci", c("conf.low", "conf.high")) |>
  rename(
    variable = group2,
    variable_level = group2_level,
    stat_name = variable_level
  ) |>
  mutate(stat_name = dplyr::recode(stat_name, "hr" = "estimate"))

# logpvalue
rtables_pval <- rtables_results |>
  filter(variable_level == "pvalue" & !is.na(stat)) |>
  select("group2_level", "stat") |>
  rename("rowname" = "group2_level")

# {cards}/{cardx}
ard_unstrat_coxph <- ard_unstrat_coxph |>
  dplyr::select(c("variable", "variable_level", "stat_name", "stat"))
attr(ard_unstrat_coxph$variable_level, "names") <- NULL
attr(ard_unstrat_coxph$stat, "names") <- NULL

# log_pval
ard_pval <- ard_pval |>
  tibble::rownames_to_column() |>
  select(c("rowname", "stat"))
ard_pval$rowname <- sub(".*vs\\.\\s*", "", ard_pval$rowname)
ard_pval$rowname <- sub("^'(.*)'$", "\\1", ard_pval$rowname)

# Compare stats
diffdf::diffdf(
  rtables_unstrat_coxph,
  ard_unstrat_coxph,
  keys = c("variable", "variable_level", "stat_name", "stat"),
  suppress_warnings = TRUE
)

# Log pvalue
diffdf::diffdf(
  rtables_pval,
  ard_pval,
  keys = c("rowname", "stat"),
  suppress_warnings = TRUE
)
```

### Survival (Timepoint)

For {rtables}: 1. Filter for event rate statistics. 2. Rename the grouping variables to match ARD 3. Apply row transformations to differentiate bounds of stats 4. Recode "pt_at_risk" to "n.risk" and "event_free_rate" to "estimate" to match ARD

For ARD: 1. Recode variable to match {rtables} 2. Format proportions to scale of 100

```{r}
# {rtables}
rtables_surv_time <- rtables_results |>
  filter(variable == "surv_6") |>
  select(c("group2_level", "variable_level", "stat")) |>
  apply_row_transformation("rate_ci", c("conf.low", "conf.high")) |>
  rename(
    group1_level = group2_level,
    stat_name = variable_level
  ) |>
  mutate(stat_name = dplyr::recode(stat_name,
    "pt_at_risk" = "n.risk",
    "event_free_rate" = "estimate"
  ))
# {cards}/{cardx}
ard_surv_time <- ard_surv_time |>
  select("group1_level", "stat_name", "stat") |>
  mutate(
    stat = case_when(
      stat_name == "estimate" ~ stat * 100,
      stat_name == "conf.low" ~ stat * 100,
      stat_name == "conf.high" ~ stat * 100,
      TRUE ~ stat
    )
  )

# Compare stats
diffdf::diffdf(
  rtables_surv_time,
  ard_surv_time,
  keys = c("group1_level", "stat_name", "stat"),
  suppress_warnings = TRUE
)
```

### Survival Diff

For {rtables}: 1. Filter for event rate statistics. 2. Rename the grouping variables to match ARD 3. Apply row transformations to differentiate bounds of stats 4. Recode "rate_diff" to "estimate" and "ztest_pval" to "p.value" to match ARD

For ARD: 1. Recode variable to match {rtables} 2. Format proportions to scale of 100

The numeric values of the results are identical but differing directions (due to the A vs. B and B vs. A type differences in the calculations).

```{r}
# {rtables}
rtables_surv_diff <- rtables_results |>
  filter(variable == "surv_diff_6" & !is.na(stat)) |>
  select(c("group2_level", "variable_level", "stat")) |>
  apply_row_transformation("rate_diff_ci", c("conf.low", "conf.high")) |>
  rename(
    group1_level = group2_level,
    stat_name = variable_level
  ) |>
  mutate(stat_name = dplyr::recode(stat_name,
    "rate_diff" = "estimate",
    "ztest_pval" = "p.value"
  ))
# {cards}/{cardx}
ard_surv_time_diff <- ard_surv_time_diff |>
  select("group1_level", "stat_name", "stat") |>
  mutate(
    stat = case_when(
      stat_name == "estimate" ~ stat * 100,
      stat_name == "conf.low" ~ stat * 100,
      stat_name == "conf.high" ~ stat * 100,
      TRUE ~ stat
    )
  )

# compare stats
diffdf::diffdf(
  rtables_surv_diff,
  ard_surv_time_diff,
  keys = c("group1_level", "stat_name", "stat"),
  suppress_warnings = TRUE
)
```
